On the presentation of orthogonal groups over polynomial rings

### 2019-05-13 dl-paper.tex
Minor fix
Abridged Lemma 19 and Corollary 3

\begin{rem} \label{rem:palpha} From the above two lemmas it follows that the value of $p_\alpha$ on an element $g\in K(A[X], M[X])$ can be computed as follows. Express $g$ as a product of elements $z_{\beta}(Xf, \xi)$ for $\beta \in \Phi$ and $c_\delta(f, X\xi)$ for some fixed $\delta \perp \alpha$. Now $p_\alpha(g)$ coincides with the sum of constant terms of the polynomials $f$ corresponding to the factors of the form $z_{-\alpha}(Xf, \xi)$ occuring in this expression. \end{rem}

@@

\begin{df} Define the map $S_\alpha(a, -) \colon P_\alpha(*) \to \St(\Phi, A[X, X^{-1}])$ on each coset $P_\alpha(m)$, $m \in M$ via the following formula:
\[ S_\alpha(a, g) = x_\alpha(aX^{-1})\cdot g \cdot x_\alpha\left(-\frac{aX^{-1}}{1 + am}\right) \cdot \{X, 1+ am\}_\alpha,\ g \in P_\alpha(m).\] \end{df}
It follows immediately from the definition of $S_\alpha(a, -)$ that for $g_1 \in P_\alpha(m)$, $g_2 \in P_\alpha(*)$ holds
\[ S_\alpha(a, g_1\cdot g_2) = S_\alpha(a, g_1) \cdot S_\alpha\left(\frac{a}{1+am}, g_2\right).\]

It is clear that the restriction of the map $S_\alpha(a, g)$ to the subgroup $P_\alpha(0)$ coincides with the conjugation map $g \mapsto {}^{x_\alpha(aX^{-1})}\!g$.

@@

Let us verify the assertion for the generator of type 4.
Choose $\beta\in \Phi$ such that $\alpha+\beta \in \Phi$. 
Substituting $s = Xf$, $\eta = X$, $\xi = aX^{-1}$ and $\alpha = -\alpha - \beta$ into~\eqref{Z5} of~\cref{Zrels} we obtain that
\begin{multline} \label{eq:Zconj3} z_{-\alpha}(X^2f, aX^{-1}) = x_{-\alpha-\beta}(\epsilon Xf) \cdot x_{-\beta}(-af) \cdot x_{\beta}(aX^2 f) \cdot x_{-\alpha}(X^2f) \cdot \\
 \cdot z_{-\alpha-\beta}(-\epsilon Xf, -\epsilon a) \cdot x_{\alpha+\beta}(-\epsilon a^2 Xf) \cdot x_{\alpha}(- a^2 f) \cdot z_{-\beta}(a f, -X), \end{multline}
where $\epsilon = N_{-\alpha-\beta,\beta}$. 
Clearly, the expression in the right-hand side is a product of generators of type 1, 2, 1, 4, 1, 1, 5, 2.

\begin{comment}
It remains to verify the assertion for the remaining generator $z_\alpha(f, X\xi)$ of type 2.
Again choose $\beta$ to be any root such that $\alpha+\beta \in \Phi$.
Substituting in~\eqref{Z5} of~\cref{Zrels} $s = -\epsilon f$, $\xi = -\epsilon \xi$, $\eta=X$ and expressing $z_\alpha(f, X\xi)$ through other terms we obtain that
\begin{multline} \nonumber z_\alpha(f, X\xi) = x_{\alpha+\beta}(\epsilon Xf) \cdot x_{\beta}(-X^2 f \xi) \cdot x_{-\beta}(f \xi) \cdot x_\alpha(f) \cdot \\ 
 \cdot z_{\alpha+\beta}(-\epsilon X f, -\epsilon \xi) \cdot z_{-\beta}(-f\xi, -X) \cdot x_{-\alpha-\beta}(-\epsilon X f \xi^2) \cdot x_{-\alpha}(-X^2 f\xi^2), \end{multline}
where $\epsilon = N_{\alpha,\beta}$.  
The factors in the right-hand side are generators of type 1, 1, 2, 2, 1, 2, 1, 4. 
It is clear, that conjugation by $x_{\alpha}(aX^{-1})$ acts trivially on the fourth factor $x_\alpha(f)$, 
while for the other factors we have already verified that their conjugates belong to $P_{\alpha}(0)$. 
\end{comment}
\end{proof}

\begin{lemma} \label{lem:Tulenbayev-formula} One has
 \[x_\alpha(aX^{-1}) \cdot x_{-\alpha}(mX) = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot 
 x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) \cdot \{1+am, X\}.\] \end{lemma}
\begin{proof} Since $\Phi$ is nonsymplectic, we can choose some $\gamma$ such that $\langle \alpha, \gamma \rangle = -1$.
 \begin{multline*}
 x_\alpha(aX^{-1}) \cdot x_{-\alpha}(mX) %= x_\alpha(X^{\langle \alpha, \gamma \rangle}a) x_{-\alpha}(X^{-\langle \alpha, \gamma \rangle}m) 
 = {}^{h_\gamma(X)}(x_\alpha(a) \cdot x_{-\alpha}(m)) = \\
 = {}^{h_\gamma(X)}\left( x_{-\alpha}\left(\frac{m}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_\alpha\left(\frac{a}{1+am}\right) \right) = \\
 = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(X^{-1}(1+am))\cdot h_\alpha(X^{-1})^{-1} \cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) = \\
 = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot \{1+am, X^{-1}\}^{-1} \cdot h_\alpha(1+am)\cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) = \\
 = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) \cdot \{1+am, X\}. \qedhere
\end{multline*}
\end{proof}

\begin{corollary} For $g \in P_\alpha(m)$ holds $S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am) \in P(\alpha, \frac{m}{1 + am})$.
Consequently one has \[ S_\alpha(a+b, g) = S_\alpha\left(b,\ S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am)\right)\cdot h_\alpha(1+am). \]\end{corollary}
\begin{proof}
Suppose $g \in P_\alpha(m)$, then $g = g_0 \cdot x_{-\alpha}(mX)$ for some $g_0 \in P_\alpha(0)$ for which we already know that $S_\alpha(a, g_0) \in P(\alpha, 0)$.
We need to verify that $h = S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am)$ lies in $P\left(\alpha, \frac{m}{1 + am}\right)$, but from Lemmas~\ref{P0_conj}--\ref{lem:Tulenbayev-formula} and the antisymmetry of Steinberg symbols we obtain that:
\begin{multline} \nonumber h = {}^{x_\alpha(aX^{-1})}\!g_0 \cdot \left(x_{\alpha}(aX^{-1}) \cdot x_{-\alpha}(mX)\right) \cdot x_{\alpha}\left(-\frac{aX^{-1}}{1+am}\right) \cdot \{X, 1+am\} \cdot h_\alpha^{-1}(1 + am) = \\
= S_\alpha(a, g_0) \cdot x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \in P_\alpha\left(\frac{m}{1+am}\right). \end{multline}
The second assertion can be verified in a straightforward manner. \end{proof}

\begin{comment}
\begin{lemma} If $\alpha - \beta \in \Phi$ then for all $g \in \St(\Phi, R[X], XM[X])$ holds $S_\alpha(a, g) \in P_\beta(0)$. \end{lemma}
\begin{proof}
It suffices to verify the assertion for the generators $z_\gamma(Xf, g)$, $\gamma \in \Phi$.
By our assumption $(\alpha, \beta) \geq 0$. 
\end{proof}
\end{comment}

@@

\begin{prop} \label{K-a-b}
 Every element $g \in K(A[X], M[X])$ can be written as a product $g_0 \cdot x_{-\alpha}(mX) \cdot x_{-\beta}(m'X)$ for some $m, m' \in M$ and $g_0 \in K(\alpha, \beta)$. Consequently, $K(\alpha, \beta)$ coincides with the intersection $K(A[X], M[X]) \cap P_\alpha(0) \cap P_\beta(0)$.
\end{prop}
\begin{proof}
 Denote by $\Sigma$ the special part of the parabolic subset $S = \Psi \cup \Phi^+$.
 Applying~\cref{thm:Stepanov} to $S$ we obtain that every element of $\St(\Phi, A[X], XM[X])$ can be presented as a product of 
  $z_\gamma(Xf, \xi)$, $\gamma \in \Sigma(S) \subseteq \Phi\setminus \Psi$ and $x_\gamma(Xf)$, $\gamma \in \Psi$.
 By~\cref{Kgen-strong} we can write $g$ as a product of these generators and generators of type 4.

@@

\begin{lemma} For $a \in A$ one has $K(\alpha, \beta)^{x_{\beta - \alpha}(a)} \subseteq K(\alpha, \beta)$. \end{lemma}
\begin{proof}
 By~\eqref{Z2}--\eqref{Z4} of~\cref{Zrels} the conjugate to the generator $g=z_\gamma(Xf, \xi)$ of type 1 by $x_{\beta-\alpha}(a)$ either coincides with $g$ or is a product of elements of type 4, 1, 1.

 Now let $g$ be a generator of type 2 or 3. It is clear that conjugation by $x_{\beta-\alpha}(a)$ either fixes $g$ or transforms it into a product of two elements of type 2 or 3.

 By~\eqref{C1}--\eqref{C3} of~\cref{Crels} the conjugate to a generator $g = c_\delta(f, X\xi)$ of type 5 either coincides with $g$ or is a product of generators of type 5, 4, 1.

 Finally, from~\cref{rem:palpha} we obtain that $z_{\alpha-\beta}(Xf, a) \in K(A[X], M[X]) \cap P_\alpha(0) \cap P_{\beta}(0) = K(\alpha, \beta)$. \end{proof}

>>>>

\begin{rem} \label{rem:palpha} From the two preceding lemmas it follows that the value of $p_\alpha$ on an element $g\in K(A[X], M[X])$ can be computed as follows. Express $g$ as a product of elements $z_{\beta}(Xf, \xi)$ for $\beta \in \Phi$ and $c_\delta(f, X\xi)$ for some fixed $\delta \perp \alpha$. Now $p_\alpha(g)$ coincides with the sum of constant terms of the polynomials $f$ corresponding to the factors of the form $z_{-\alpha}(Xf, \xi)$ occuring in this expression. \end{rem}

@@

\begin{df} Define the map $S_\alpha(a, -) \colon P_\alpha(*) \to \St(\Phi, A[X, X^{-1}])$ on each coset $P_\alpha(m)$, $m \in M$ via the following formula:
\[ S_\alpha(a, g) = x_\alpha(aX^{-1})\cdot g \cdot x_\alpha\left(-\frac{aX^{-1}}{1 + am}\right) \cdot \{X, 1+ am\}_\alpha,\ g \in P_\alpha(m).\] \end{df}
It follows immediately from the definition of $S_\alpha(a, -)$ that for $g_1 \in P_\alpha(m)$, $g_2 \in P_\alpha(*)$ holds
\begin{equation} \label{eq:Smult} S_\alpha(a, g_1\cdot g_2) = S_\alpha(a, g_1) \cdot S_\alpha\left(\frac{a}{1+am}, g_2\right).\end{equation}

It is clear that the restriction of the map $S_\alpha(a, g)$ to the subgroup $P_\alpha(0)$ coincides with the map of left conjugation by $x_\alpha(aX^{-1})$.

@@

Let us verify the assertion for the generator of type 4.
Choose $\beta\in \Phi$ such that $\alpha+\beta \in \Phi$. 
Substituting $s = Xf$, $\eta = X$, $\xi = aX^{-1}$ and $\alpha = -\alpha - \beta$ into~\eqref{Z5} of~\cref{Zrels} we obtain that
\begin{multline} \label{eq:Zconj3} z_{-\alpha}(X^2f, aX^{-1}) = x_{-\alpha-\beta}(\epsilon Xf) \cdot x_{-\beta}(-af) \cdot x_{\beta}(aX^2 f) \cdot x_{-\alpha}(X^2f) \cdot \\
 \cdot z_{-\alpha-\beta}(-\epsilon Xf, -\epsilon a) \cdot x_{\alpha+\beta}(-\epsilon a^2 Xf) \cdot x_{\alpha}(- a^2 f) \cdot z_{-\beta}(a f, -X), \end{multline}
where $\epsilon = N_{-\alpha-\beta,\beta}$. 
Clearly, the expression in the right-hand side is a product of generators of type 1, 2, 1, 4, 1, 1, 5, 2. \end{proof}

\begin{lemma} \label{lem:Tulenbayev-formula} 
One has 
$ S_\alpha(a, x_{-\alpha}(mX)) = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am).$
\end{lemma}
\begin{proof} Since $\Phi$ is nonsymplectic, we can choose some $\gamma$ such that $\langle \alpha, \gamma \rangle = -1$.
 \begin{multline*}
 x_\alpha(aX^{-1}) \cdot x_{-\alpha}(mX) %= x_\alpha(X^{\langle \alpha, \gamma \rangle}a) x_{-\alpha}(X^{-\langle \alpha, \gamma \rangle}m) 
 = {}^{h_\gamma(X)}(x_\alpha(a) \cdot x_{-\alpha}(m)) = \\
 = {}^{h_\gamma(X)}\left( x_{-\alpha}\left(\frac{m}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_\alpha\left(\frac{a}{1+am}\right) \right) = \\
 = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(X^{-1}(1+am))\cdot h_\alpha(X^{-1})^{-1} \cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) = \\
 = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot \{1+am, X^{-1}\}^{-1} \cdot h_\alpha(1+am)\cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) = \\
 = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) \cdot \{1+am, X\},
\end{multline*}
which implies the assertion.
\end{proof}

\begin{corollary} For $g \in P_\alpha(m)$ holds $S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am) \in P(\alpha, \frac{m}{1 + am})$.
Consequently one has \[ S_\alpha(a+b, g) = S_\alpha\left(b,\ S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am)\right)\cdot h_\alpha(1+am). \]\end{corollary}
\begin{proof}
Fix an element $g \in P_\alpha(m)$ and write it $g = g_0 \cdot x_{-\alpha}(mX)$ for some $g_0 \in P_\alpha(0)$.
Now by~\eqref{eq:Smult} and~\cref{P0_conj} one has
\begin{multline} \nonumber
S_\alpha(a, g) \cdot h_{\alpha}^{-1}(1+am) = S_\alpha(a, g_0) \cdot S_\alpha(a, x_{-\alpha}(mX)) \cdot h_{\alpha}^{-1}(1+am) = 
\\ = S_\alpha(a, g_0) \cdot x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot\langle -a, m \rangle \in P_\alpha\left(\frac{m}{1+am}\right).
\end{multline}

The second assertion can be verified in a straightforward manner. \end{proof}

@@

\begin{prop} \label{K-a-b}
 Every element $g \in K(A[X], M[X])$ can be written as a product $g_0 \cdot x_{-\alpha}(mX) \cdot x_{-\beta}(m'X)$ for some $m, m' \in M$ and $g_0 \in K(\alpha, \beta)$. Consequently, $K(\alpha, \beta)$ coincides with the intersection $K(A[X], M[X]) \cap P_\alpha(0) \cap P_\beta(0)$.
\end{prop}
\begin{proof}
 Denote by $\Sigma$ the special part of the parabolic subset $S = \Psi \cup \Phi^+$.
 Applying~\cref{thm:Stepanov} to $S$ we obtain that every element of $\St(\Phi, A[X], XM[X])$ can be presented as a product of 
  $z_\gamma(Xf, \xi)$, $\gamma \in \Sigma(S) \subseteq \Phi\setminus \Psi$ and $x_\gamma(Xf)$, $\gamma \in \Psi$.
 By~\cref{Kgen-strong} we can write $g$ as a product of these generators and generators of type 5.

@@

\begin{lemma} For $a \in A$ one has $K(\alpha, \beta)^{x_{\beta - \alpha}(a)} \subseteq K(\alpha, \beta)$. \end{lemma}
\begin{proof}
 By~\eqref{Z2}--\eqref{Z4} of~\cref{Zrels} the conjugate to the generator $g=z_\gamma(Xf, \xi)$ of type 1 by $x_{\beta-\alpha}(a)$ either coincides with $g$ or is a product of elements of type 4, 1, 1. Now if $g = x_\epsilon(Xf)$, $\epsilon \in \{\pm \alpha, \pm \beta\}$ is a generator of type 2 or 3, the conjugation by $x_{\beta-\alpha}(a)$ either fixes $g$ or transforms it into a product of either two elements of type 2 or two elements of type 3. By~\eqref{C1}--\eqref{C3} of~\cref{Crels} the conjugate of a generator $g = c_\delta(f, X\xi)$ of type 5 either coincides with $g$ or is a product of generators of type 5, 4, 1. Finally, from~\cref{rem:palpha} we obtain that $z_{\alpha-\beta}(Xf, a) \in K(A[X], M[X]) \cap P_\alpha(0) \cap P_{\beta}(0) = K(\alpha, \beta)$. \end{proof}
